Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms

### 2017-08-15 paper/paper.tex
Patrick's comments

The clique and $k{\downarrow}$ algorithms already make use of fine-granularity bit-parallelism. To
introduce more coarse-grained thread parallelism, we will parallelise the search process: viewing
backtracking search as forming a tree, we can explore different portions of the tree using different
threads. We use a shared incumbent, so better solutions found by one thread can be used others
immediately. This technique has a long history \citep{o:BaderHC05}. Of particular
interest to us are so-called \emph{anomalies}
\citep{DBLP:journals/cacm/LaiS84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95}:
because we are not performing a fixed amount of work, we should have no expectation of a linear
speedup, and instead we could see a sublinear speedup (much less than $n$ from $n$ processors, if
speculative work turns out to be wasted) or a superlinear speedup (much more than $n$ from $n$
processors, if a strong incumbent is found more quickly). An absolute slowdown (a speedup much less
than 1) is also possible under some models.

@@

We will use this mechanism for our experiments.  The clique-based maximum common subgraph algorithm
effectively differs only in the preprocessing stage, and the clique-inspired connected algorithm
described by \citet{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that may be parallelised in
exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
rather than the original three, since maximum common subgraph instances appear to give even more
irregular search trees than normal clique problem instances.

@@

In principle the McSplit algorithm may be parallelised in exactly the same way. However, this
algorithm makes heavy use of an in-place, backtrackable data structure, which is not copied for
recursive calls. In order to introduce the \emph{potential} for parallelism, we must make
copies of the state data structure. Implemented na{\"\i}vely, this can give an order of magnitude
slowdown to the sequential algorithm, which can be hard to recover using parallelism. To lessen the
effects, rather than copying state for each recursive call, we copy once before the main branching
loop, and then copy that copy in each ``helper'' thread, replaying the branching loop without making
duplicate recursive calls.  (We are not convinced that a better approach is not possible\ldots)

>>>>

The clique and $k{\downarrow}$ algorithms already make use of fine-granularity bit-parallelism. To
introduce more coarse-grained thread parallelism, we will parallelise the search process: viewing
backtracking search as forming a tree, we can explore different portions of the tree using different
threads. We use a shared incumbent, so better solutions found by one thread can be used by others
immediately. This technique has a long history \citep{o:BaderHC05}. Of particular
interest to us are so-called \emph{anomalies}
\citep{DBLP:journals/cacm/LaiS84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95}:
because we are not performing a fixed amount of work, we should have no expectation of a linear
speedup, and instead we could see a sublinear speedup (much less than $n$ from $n$ processors, if
speculative work turns out to be wasted) or a superlinear speedup (much more than $n$ from $n$
processors, if a strong incumbent is found more quickly). An absolute slowdown (a speedup much less
than 1) is also possible under some models.

@@

We will use this mechanism for our experiments.  The clique-based maximum common subgraph algorithm
effectively differs only in the preprocessing stage, and the clique-inspired connected algorithm
described by \citet{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that it may be parallelised in
exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
rather than the original three, since maximum common subgraph instances appear to give even more
irregular search trees than normal clique problem instances.

@@

In principle the McSplit algorithm may be parallelised in exactly the same way. However, this
algorithm makes heavy use of an in-place, backtrackable data structure, which is not copied for
recursive calls. In order to introduce the \emph{potential} for parallelism, we must make
copies of the state data structure. Implemented na{\"\i}vely, this can give an order of magnitude
slowdown to the sequential algorithm, which can be hard to recover using parallelism. To lessen the
effects, rather than copying state for each recursive call, we copy once before the main branching
loop, and then copy that copy in each ``helper'' thread, replaying the branching loop without making
duplicate recursive calls.  (We believe a better approach using partial recomputation may be
possible, and intend to investigate this further in the future.)
